---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Markov Chain Monte Carlooo

```{r}
library(tidyverse)
library(brms)
```

The king visiting islands proportional to their population.

```{r}
num_weeks <- 1e5
positions <- rep(0, num_weeks)
current <- 7

for (i in 1:num_weeks) {
  # record my current position for plotting mainly.
  positions[i] <- current
  
  # proposal island position
  proposal <- current + sample(c(-1,1), size = 1)
  
  # loop around
  if (proposal < 1) proposal <- 10
  if (proposal > 10) proposal <- 1
  
  # move or not
  prob_move <- proposal / current
  current <- ifelse(runif(1) < prob_move, proposal, current)
}

plot(positions[1:100])
plot(table(positions))
```

Important to keep this in mind for intuition:

The “islands” in our objective are parameter values, and they need not be discrete, but can instead take on a continuous range of values as usual.

The “population sizes” in our objective are the posterior probabilities at each parameter value.

The “weeks” in our objective are samples taken from the joint posterior of the parameters in the model.

A new fancy term to think about : Concentration of Meausre.
