---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r}
library(brms)
library(tidyverse)

library(wesanderson)

theme_set(
  theme_minimal() + 
    theme(
      panel.grid = element_blank()
    )
)
```

Something to keep in mind:

The essential problem is this: When what we wish to predict is a count, the scale of the parameters is never the same as the scale of the
outcome.


## Pro and Anti-Social Chimpanzees.

```{r}
data(chimpanzees, package = "rethinking")
d <- chimpanzees
rm(chimpanzees)

glimpse(d)
```


The Index Variable approach:


```{r}
d <- d %>% mutate(
  treatment = factor(1 + prosoc_left + 2 * condition)
) %>% mutate(
  labels = factor(treatment,
                  levels = 1:4,
                  labels = c("r/n", "l/n", "r/p", "l/p"))
)
# Just looking at general summary:
d %>% group_by(treatment) %>% summarize(mean(pulled_left))
```


Let's fit an intercept only model and compare priors:
Important bit on thinking about priors for logit link models:

A flat prior in the logit space is not a flat prior in the outcome probability space.

```{r}
b11.1 <- brm(
  data = d,
  family = binomial,
  pulled_left | trials(1) ~ 1,
  
  prior = c(
    prior(normal(0, 10), class = Intercept)
  ),
  iter = 2000, warmup = 1000, chains = 4, cores = 4,
  sample_prior = T,
)
```

A much more sensible prior:

```{r}
b11.1b <- brm(
  data = d,
  family = binomial,
  pulled_left | trials(1) ~ 1,
  
  prior = c(
    prior(normal(0, 1.5), class = Intercept)
  ),
  iter = 2000, warmup = 1000, chains = 4, cores = 4,
  sample_prior = T,
)

```


The prior for p_i which is equal to exp(a)/ 1 + exp(a) looks something like this.

```{r}
curve(exp(x)/ (1 + exp(x)), from = -4, to = 4)
```

Let's compare the priors.

```{r}
bind_rows(
  prior_draws(b11.1),
  prior_draws(b11.1b)
) %>% 
  mutate(p = inv_logit_scaled(Intercept),
         w = factor( rep(c(10, 1.5), each = n()/2), levels = c(10, 1.5))
         ) %>% 
  ggplot(aes(x = p, fill = w))+
  geom_density(alpha = 0.3, adjust = 0.1)+
  scale_fill_manual(
    values = c("10" = "red", "1.5" = "yellow")
  )


```

Some todos from brms and tidyverse:
[] Model pulled_left with only varitions in treatment
[] Model pulled_left including both actors and treatment, and see their coefficient plots.
[] Computing and Understanding the differences:

For left hand side
No partner - Partner = Expectation is (low odds - high odds) of pulling left lever. We should expect high magnitude negative.


For Right hand side
No partner - Partner = Expectation is (high odds - low odds) of pulling left lever
                     = Expectations is (low odds - high odds) of pulling right lever
                     
We should expect high positive magnitude. 


# Model pulled_left with one intercept and variations in treatment:

This verbose approach is only required if we want separate columns for each treatment values.


```{r}
b11.3 <- brm(
  data = d,
  family = binomial,
  bf(pulled_left | trials(1) ~ a + b,
     a ~ 1,
     b ~ 0 + treatment,
     nl = T),
  prior = c(
    prior(normal(0, 1.5), nlpar = a),
    prior(normal(0, 0.5), nlpar = b, coef = treatment1),
    prior(normal(0, 0.5), nlpar = b, coef = treatment2),
    prior(normal(0, 0.5), nlpar = b, coef = treatment3),
    prior(normal(0, 0.5), nlpar = b, coef = treatment4)
  ),
  
  iter = 2000, warmup = 1000, chains = 4, cores = 4,
  file = "brms_code/fits/b11.03"
)


```

Non Verbose Approach:

```{r}
b11.3a <- brm(
  data = d,
  family = binomial(),
  bf(pulled_left | trials(1) ~ a + b,
     a ~ 1,
     b ~ 0 + treatment, 
     nl = T),
  
  prior = c(
    prior(normal(0, 1.5), nlpar = a),
    prior(normal(0, 0.5), nlpar = b)
  ),
  
  iter = 2e3, warmup = 1e3, cores = 4, chains = 4,
  file = "brms_code/fits/b11.03a"
)
```


# Among left-handers, what's the difference between no partner and partner?

```{r}
as_draws_df(b11.3) %>% 
  mutate(diff24 = b_b_treatment2 - b_b_treatment4) %>% 
  ggplot(aes(x = diff24))+
  geom_density(fill = "skyblue")+
  geom_vline(xintercept = 0, linetype = 2)
```

The answer is not a lot.

# Incorporating the variation for individual monkeys as well:

```{r}
d <- d %>% mutate(actor = factor(actor))

b11.4 <- brm(
  data = d,
  family = binomial,
  bf(pulled_left | trials(1) ~ a + b,
     a ~ 0 + actor,
     b ~ 0 + treatment, 
     nl = T),
  prior = c(
    prior(normal(0, 1.5), nlpar = a),
    prior(normal(0, 0.5), nlpar = b)
  ),
  
  iter = 2000, warmup = 1000, chains = 4, cores = 4,
  file = "brms_code/fits/b11.04"
)
```

# Let's make a coefficient plot for the treatment:

```{r}
library(tidybayes)

post <- as_draws_df(b11.4); glimpse(post)

post %>% 
  pivot_longer(contains("treatment")) %>% 
  mutate(probability = inv_logit_scaled(value),
         treatment = str_remove(name, "b_b_treatment")) %>% 
  ggplot(aes(x = probability, y = treatment))+
  stat_pointinterval(.width = .95, linewidth = .5)
```


# Let's make a contrast coefficient plot for differences:

```{r}
post %>% mutate(
  db13 = b_b_treatment1 - b_b_treatment3,
  db24 = b_b_treatment2 - b_b_treatment4
) %>% pivot_longer(db13:db24) %>% 
  mutate(diffs = factor(name, levels = c("db24", "db13"))) %>% 
  
  ggplot(aes(x = value, y = diffs))+
  stat_pointinterval(.width = .95, linewidth = .5)+
  geom_vline(xintercept = 0, linetype = 2)
```


### Relative change in odds.

I want to know the change in odds of pulling the left lever if I go from treatment 2 (left handed: no partner) to treatment 4 (left handed: a partner)

```{r}
post %>% 
  mutate(proportional_odds = exp(b_b_treatment4 - b_b_treatment2)) %>% 
  mean_qi(proportional_odds)
```

The answer is 0.931 i.e. the odds of pulling the left lever decrease by 8%.

### Using condensed binomial counts.

```{r}
d <- d %>% mutate(
  side = factor(prosoc_left + 1), # right is 1, left is 2
  cond = factor(condition + 1 ) # no partner is 1, partner is 2
)
```

Aggregating the data.

```{r}
d_aggregate <- d %>% 
  group_by(treatment, actor, side, cond) %>% 
  summarize(left_pulls = sum(pulled_left))

```

The aggregated model:

```{r}
b11.6 <- brm(
  data = d_aggregate,
  family = binomial,
  bf(left_pulls | trials(18) ~ a + b,
     a ~ 0 + actor,
     b ~ 0 + treatment,
     nl  = T),
  prior = c(
    prior(normal(0, 1.5), nlpar = a),
    prior(normal(0, 0.5), nlpar = b)
  ),
  
  iter = 2e3, warmup = 1e3, chains = 4, cores = 4,
  sample_prior = T,
  file = "brms_code/fits/b11.06"
)
```

Practising drawing coefficient plots.

```{r}
as_draws_df(b11.6) %>% 
  select(contains("treatment")) %>% 
  pivot_longer(everything()) %>% 
  mutate(treatment = str_remove(name, "b_b_treatment"),
         probability = inv_logit_scaled(value)) %>% 
  ggplot(aes(x = probability, y = treatment))+
  stat_pointinterval(.width = 95, linewidth = 1)
```



